Derive final annotations based on NBAtlas reference

### 2025-08-25 analyses/cell-type-neuroblastoma-04/final-annotation.Rmd
add more NE annotations

1. If `SingleR` and `scANVI/scArches` labels exactly agree, we assign that label
2. If broad label families (e.g., `T cell` would be the broad family for `CD4+ T cell`) for `SingleR` and `scANVI/scArches` labels agree, we assign the broad family label.
  * See [`references/nbatlas-label-map.tsv`](references/nbatlas-label-map.tsv) and [`references/README.md`](references/README.md) for how labels were mapped to families.
3. If `SingleR` and `scANVI/scArches` labels disagree, we then compare to the consensus cell type; if at least one inference agrees with the consensus cell type, we assign the inferred label.
  * We perform this first at the label level and then at the broad family level.
4. Finally, following the conclusions of Bonine et al. (2024), we designate any cell labeled as `Neuroendocrine` as a tumor cell.

@@

@@

```{r}
final_annotation_df <- annotation_df |> 
  dplyr::mutate(
    # For all comparisons, make sure we aren't comparing NA to NA and getting TRUE
    final_label = dplyr::case_when(
      ########### Check for exact match between SingleR/scANVI
      !is.na(scanvi_label_ontology) & singler_label_ontology == scanvi_label_ontology ~ scanvi_label, 
      ########### Check for family match between SingleR/scANVI
      !is.na(scanvi_family_ontology) & singler_family_ontology == scanvi_family_ontology ~ scanvi_family,
      ########## Now use agreement with consensus to assign a label: first by label, and then by family
      !is.na(consensus_ontology) & singler_label_ontology  == consensus_ontology ~ singler_label,
      !is.na(consensus_ontology) & scanvi_label_ontology   == consensus_ontology ~ scanvi_label,
      !is.na(consensus_family_ontology) & singler_family_ontology == consensus_family_ontology ~ singler_family,
      !is.na(consensus_family_ontology) & scanvi_family_ontology  == consensus_family_ontology ~ scanvi_family,
      # Everything else is Unknown
      .default = "Unknown"
    )
  ) |>
  # Now, we can bring the final ontologies into the data frame
  dplyr::inner_join(
    ontology_df |> dplyr::rename(final_label = NBAtlas_label, final_id = CL_ontology_id), 
    by = "final_label"
  ) |>
  # add tumor column for visualization
  dplyr::mutate(cell_class = dplyr::case_when(
    final_label == "Neuroendocrine" ~ "tumor", 
    final_label == "Unknown" ~ "unknown", 
    .default = "normal"
  ))
```

>>>>

1. If `SingleR` and `scANVI/scArches` labels exactly agree, we assign that label
2. If broad label families (e.g., `T cell` would be the broad family for `CD4+ T cell`) for `SingleR` and `scANVI/scArches` labels agree, we assign the broad family label.
    * See [`references/nbatlas-label-map.tsv`](references/nbatlas-label-map.tsv) and [`references/README.md`](references/README.md) for how labels were mapped to families.
3. If `SingleR` and `scANVI/scArches` labels disagree, we then compare to the consensus cell type; if at least one inference agrees with the consensus cell type, we assign the inferred label.
    * We perform this first at the label level and then at the broad family level.
4. If the consensus cell type is `Unknown` and one of `SingleR` or `scANVI/scArches` labels is `Neuroendocrine` and the other label is one of `Schwann cell`, `Stromal other` or `fibroblast`, we assign the `Neuroendocrine` label
5. Finally, following the conclusions of Bonine et al. (2024), we designate any cell labeled as `Neuroendocrine` as a tumor cell.

@@

# Define vector of tumor-like cell types
tumorlike_cells <- c("Schwann", "Stromal other", "Fibroblast")

# pull out NE ontology for assignment code
ne_ontology <- ontology_slim_df$CL_ontology_id[ontology_slim_df$NBAtlas_label == "Neuroendocrine"]

@@

```{r}
final_annotation_df <- annotation_df |> 
  dplyr::mutate(
    # For all comparisons, make sure we aren't comparing NA to NA and getting TRUE
    final_label = dplyr::case_when(
      ########### Check for exact match between SingleR/scANVI
      !is.na(scanvi_label_ontology) & singler_label_ontology == scanvi_label_ontology ~ scanvi_label, 
      ########### Check for family match between SingleR/scANVI
      !is.na(scanvi_family_ontology) & singler_family_ontology == scanvi_family_ontology ~ scanvi_family,
      ########## Now use agreement with consensus to assign a label: first by label, and then by family
      !is.na(consensus_ontology) & singler_label_ontology  == consensus_ontology ~ singler_label,
      !is.na(consensus_ontology) & scanvi_label_ontology   == consensus_ontology ~ scanvi_label,
      !is.na(consensus_family_ontology) & singler_family_ontology == consensus_family_ontology ~ singler_family,
      !is.na(consensus_family_ontology) & scanvi_family_ontology  == consensus_family_ontology ~ scanvi_family,
      ########## Assign Neuroendocrine where possible; for this step, we refer to NBAtlas labels directly since
      # the `tumorlike_cells` don't all have ontology ids
      is.na(consensus_ontology) & singler_label == "Neuroendocrine" & scanvi_label %in% tumorlike_cells ~ "Neuroendocrine",
      is.na(consensus_ontology) & scanvi_label == "Neuroendocrine" & singler_label %in% tumorlike_cells ~ "Neuroendocrine",
      # Everything else is Unknown
      .default = "Unknown"
    )
  ) |>
  # Now, we can bring the final ontologies into the data frame
  dplyr::inner_join(
    ontology_df |> dplyr::rename(final_label = NBAtlas_label, final_id = CL_ontology_id), 
    by = "final_label"
  ) |>
  # add tumor column for visualization
  dplyr::mutate(cell_class = dplyr::case_when(
    final_label == "Neuroendocrine" ~ "tumor", 
    final_label == "Unknown" ~ "unknown", 
    .default = "normal"
  ))
```
